Water Dynamics in Fogera and the Upper Blue Nile\\Farmers perspectives and remote sensing\bigskip

### 2015-03-30 grassposter/2015_EGU_WD_Ethiopia/poster.tex
Finalised poster
git-svn-id: https://svn.osgeo.org/grass/grass-promo@64937 15284696-431f-4ddb-bdfa-cd5b030d7da7

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{Spring Maize and food security}{
We found that in one case (figure 3), farmers perspective on the Spring maize crop sensibility to variability of rainfall can be quantified in space and time by remote sensing cumulative transpiration. A crop transpiration gap of 1-2.5 mm/day for about 20 days is to be overcome, a full amount of 20 to 50 mm, depending on the type of year deficit. Such gap can be overcome, even by temporary supplemental irrigation practices.\newline\linebreak
We argue that a better understanding of the dynamics of water variability and the implications on water management, agricultural production and livelihoods should consider small-holder farmers’ perceptions, concerns and responses. This is essential to devise well-targeted policies and interventions that can improve livelihoods, as well as optimising water management and agricultural production.

\begin{center}
 \includegraphics[width=0.65\textwidth]{./images/fig1}\\
 \vspace{5mm}
 Figure 3: Transpiration deficiency, month of May, one year out four.
\end{center}

Cumulative transpiration on a yearly basis highlights different years, and heterogeneity periods (breaks) in the growing curve. Looking closely into the May-July period, one can identified the transpiration gap found most of the time for the early Maize crop much discussed by farmers. Out of 14 years, only 4 seem to have made the transpiration leap above the others (red ellipse in Figure 3). The leap can be quantified by an average of 20mm of rainfall over a period of 20 days, 1 mm of transpiration per day is missing at minimum (red ellipse) and 2.5mm/day at maximum (yellow ellipse) from mid-June to early July.

@@

%\smallskip
%\hrulefill
%\vspace{-5pt}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{Yearly Transpiration gap in the Upper Nile Basin}{
\smallskip

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{Conclusions}{
\smallskip
Space science (i.e. remote sensing plant transpiration data and analysis) complements and augments farmer’s responses to rainfall variability. In such marginal landscapes access to this information can mean the difference between seasonal hunger and food security.
}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{Spring Maize and food security}{
\begin{center}
	\begin{tabular}{ll}
 	\begin{tabular}{p{0.35\textwidth}}
Looking into the cumulative transpiration yearly signals (Figure 3), there is non uniformity, thus variations in water availability to vegetation. Following farmers' interviews and looking more closely into the April-June period, we found that in one case (figure 4), farmers perspective on the Spring maize crop sensibility to variability of rainfall can be quantified in space and time by remote sensing cumulative transpiration. A crop transpiration gap of 1-2.5 mm/day for about 20 days is to be overcome, a full amount of 20 to 50 mm, depending on the type of year deficit. Such gap can be overcome, even by temporary supplemental irrigation practices.\newline\linebreak
We argue that a better understanding of the dynamics of water variability and the implications on water management, agricultural production and livelihoods should consider small-holder farmers’ perceptions, concerns and responses. This is essential to devise well-targeted policies and interventions that can improve livelihoods, as well as optimising water management and agricultural production.\newline\linebreak
Cumulative transpiration on a yearly basis highlights different years, and heterogeneity periods (breaks) in the growing curve. Looking closely into the May-July period, one can identified the transpiration gap found most of the time for the early Maize crop much discussed by farmers. Out of 14 years, only 4 seem to have made the transpiration leap above the others (red ellipse in Figure 4). The leap can be quantified by an average of 20mm of rainfall over a period of 20 days, 1 mm of transpiration per day is missing at minimum (red ellipse) and 2.5mm/day at maximum (yellow ellipse) from mid-June to early July.
	\end{tabular}
 	& 
 	\begin{tabular}{c}
 	\includegraphics[width=0.55\textwidth]{./images/fig1big}\\
 	\includegraphics[width=0.55\textwidth]{./images/fig1}\\
	\end{tabular}
	\end{tabular}
	\vspace{5mm}
\end{center}
\begin{flushright}
Figure 3\&4: Cumulative transpiration per year \& month of May.\\
\end{flushright}

@@

\smallskip
\hrulefill
%\vspace{-5pt}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{Yearly Transpiration gap in the Upper Nile Basin}{
\smallskip
The Upper Nile Basin has a dramatic range of altitude from 483m reaching to 4517m at most. While most of the high altitude areas and various water bodies (in blue in Figure 4) are not or very less vegetated, the transpiration gap is relatively uninteresting. On the other hand, very clearly (in red in Figure 4), valleys are under deficit of transpiration from their potential level, vegetation is under stress.

@@

This averaged daily deficit for a full year does not do justice to seasonal variations, but encompasses the potential for analysis of much longer temporal periodicity, we are only starting to decifer the monthly and yearly structures. It is clear at this point that Fogera woreda and one of its basin Gumura are a more common case than thought initially in terms of yearly transpiration gap budget.\newline

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{Conclusions}{
\smallskip
Space science as in remote sensing plant transpiration data and analysis, enables to quantify the lack of water consumption to reach a minimum Spring maize harvest. This knowledge potentially complements and augments farmer’s resilience and responsiveness to rainfall variability. In such marginal landscapes access to this information, and development of a coping mechanism, can mean the difference between seasonal hunger and food security.\newline\linebreak

Extending the exploration field of view to the Upper Nile Basin, we can see the Fogera woreda and its basin Gumura seem to have a fairly common yearly behaviour for watersheds. The lower section of the valleys seem to be consistently to have vegetation under water stress. Searching for monthly periodicity in the gap of transpiration is a subject of further studies.\newline
